---
title: "Eating Out"
author: "Aeilani Felix"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2024"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## There were two years in which the General Social Survey asked about the frequency of eating out. The variable name you are looking for is "eatout".
### Find those years and for the first year the question was asked download that years data in a stata format - its name will be "GSSxxxx.dta" where the xxxx is replaced by the year.  For example if the year was 1983 its name would be GSS1983.dta. Your assignment is to download the GSS yearly dataset for the first year the question was asked and then  write up the answer to the question by editing the  code block in line 31 to eval=TRUE and finish the job. 

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
```
```{r, eval=TRUE, include=F}
GSS <-read.dta(file="GSS1996.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```


# Methods 
We will be using our graphical, Numerical, and Inferential results to figure out if there is no difference among the various races with frequency of eating out, and if there is a statistical difference between at least two races with the frequency of eating out.

# Graphical 

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```

This graph shows us that there are about 610 white people who eat out often and about 150 white that don't eat out. There are about 95 black people who eat out and about 40 black people who don't go eat out. About 40 others go and eat out and about 10 don't go eat out. In other words 85% of white people go eat out and 15% of white people don't go eat out, for black people about 70% go and eat out and about 25% don't go eat out, for the others about 75% go and eat out and about 20% don't go eat out.

# Numerical 

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```

There are 83% of white people go eat out and 17% who don't go out and eat, For black people there are 69% who do go out and eat and there are 31% who don't go eat out, For the others there are 78% who do go out to eat and there are 22% who don't go out to eat.

# Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
```{r}
fisherXY(mydata$race,mydata$eatout, printAll = TRUE)
```

The fisher test clearly shows us that the only thing that is statistically different was the data about black and white the p value was to small so we have to reject the HO any data that involves the others variable the p value was to high so they were very similar so we must reject the $H_0$.

# Discussion and Conclusion

In conclusion there is a difference between the white and black with frequency of eating out there are a statistical difference between at least two races with the frequency of eating out the only variables that are different are the White and the Black data there p value is to low so we have to reject the $H_0$.